---
title: "R4DSsec3-9"
author: "Sam Xu"
date: "April 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
library(mapproj)
```

## Summary
  For ggplot, coordinates can be adjusted in purose. Below are some situations where you might want to make changes.
  
### To switches the x and y axes, use coord_flip():
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```
  
### While plotting spatial data, use coord_quickmap() to sets the aspect ratio correctly for maps:
```{r}
nz <- map_data("nz")
ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")
ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```
  
### To use polar coordinates, coord_polar():
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```
  

## Exercise solution
### 1.
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = 1, fill = clarity), position = "fill") +
  coord_polar(theta = "y")
```
  
### 2.
  Labs function can modify axis, legend, and plot labels. For example:  
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = 1, fill = clarity), position = "fill")
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = 1, fill = clarity), position = "fill") +
  labs(title = "Stack plot", subtitle = "Count")
```
  
### 3.
  They do similar jobs. Map projections in 'coord_map' require package 'mapproj' and they do not, in general, preserve straight lines, so this requires considerable computation. 'coord_quickmap' is a quick approximation that does preserve straight lines. It works best for smaller areas closer to the equator.  

### 4.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```
  
  The plot shows that hwy and cty are highly correlated and the relationship is linear.  
  'coord_fixed' forces a specified ratio between the physical representation of data units on the axes. Here by default the ratio is 1. It ensures that one unit on the x-axis is the same length as one unit on the y-axis. In some sense it avoided a misleading message that correlation between hwy and cty was small.  
  'geom_abline' added a diagonal reference line to the plot.
















